---
title: "F6 interactome project: Monocle3 trajectory analysis"
author: "Jen-Chien Chang"
date: "`r format(Sys.time(), '%Y/%m/%d')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, cache=FALSE, fig.show='hold', fig.align="center")
```

---

# Note

* 

---

# Basic setup

<span style="color:red">template code</span> <a href="#top">Back to top</a>
```{r}

```


Load R library
```{r warning=FALSE, results='hide'}
# Data manipulation
library(dplyr)
library(magrittr)
library(reshape2)

# Plot
library(ggplot2)
library(patchwork)
library(ggpubr)
library(rstatix)
library(pheatmap)
#library(ggfortify)    # for plot PCA with ggplot2 autoplot 

# File I/O
#library(readxl)

# Seurat
library(Seurat)
#library(SeuratDisk)

# monocle
library(monocle3)
library(SeuratWrappers)

# Signac
#set.seed(2021)
#library(Signac)
#library(GenomeInfoDb)
#library(EnsDb.Mmusculus.v79)
#library(GenomicRanges)

# Multiprocessing
#library(future)
#plan("multiprocess", workers = 12)
#plan("sequential")

# Math
#library(usedist)

# For AUCell
#library(AUCell)
#library(GSEABase)

# sceasy (seurat <-> annData )
#library(sceasy)
#library(anndata) #directly deal with annData in R

```


Global variables
```{r}
# data directory
osc_fs_home_base <- "../../../../../osc-fs_home"
home_base <- "../../../../../home"

# color setting 
sampleID_color <- c("#e31a1c","#fb9a99","#33a02c","#b2df8a","#1f78b4","#a6cee3")
sampleType_color <- c("#e31a1c","#33a02c","#1f78b4")

# ggplot2 theme setting
theme_set(theme_bw()) 
```


---

# Trajectory analysis using scRNA-seq data

Load processed scRNA-seq data
```{r}
All.rna.flt <- readRDS("data/20220331_All.rna.flt.rds")
  
```


```{r}
cds <- as.cell_data_set(All.rna.flt)
cds <- cluster_cells(cds)
```


```{r}
plot_cells(cds, show_trajectory_graph = FALSE)
plot_cells(cds, color_cells_by = "partition", show_trajectory_graph = FALSE)
```

Learn trajectories
```{r}
cds <- learn_graph(cds)
```


```{r}
plot_cells(cds, label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = FALSE)
plot_cells(cds, label_groups_by_cluster = FALSE, label_leaves = TRUE, label_branch_points = FALSE)
plot_cells(cds, label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = TRUE)
plot_cells(cds, label_groups_by_cluster = FALSE, label_leaves = TRUE, label_branch_points = TRUE)
```


Calculate pseudotime along the trajectory

---

## Use harmony-integrated latent space

```{r}
All.rna.flt.dup <- All.rna.flt

All.rna.flt.dup@reductions[["umap"]] <- All.rna.flt.dup@reductions[["umap_harmony"]]
All.rna.flt.dup@reductions[["pca"]] <- All.rna.flt.dup@reductions[["harmony"]]

cds2 <- as.cell_data_set(All.rna.flt.dup)
cds2 <- cluster_cells(cds2)
plot_cells(cds2, color_cells_by = "partition", show_trajectory_graph = FALSE)

# Learn graph but don't use partition
cds2 <- learn_graph(cds2, use_partition = FALSE)
plot_cells(cds2, label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = FALSE)

# Deault use partition
#cds2 <- learn_graph(cds2)
#plot_cells(cds2, label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = FALSE)

```


Don't use partition
```{r}
cds2 <- learn_graph(cds2, use_partition = FALSE)
plot_cells(cds2, label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = FALSE)
```






```{r}
cds2 <- order_cells(cds2)
```


```{r}
plot_cells(cds2, color_cells_by = "pseudotime", label_cell_groups = FALSE, label_leaves = FALSE, label_branch_points = FALSE)
```

Extract pseudotime for each cell
```{r}
principal_graph_aux(cds2)$UMAP$pseudotime %>% head
```



Find DE along pseudotime
```{r}
# cannot parallel on windows, 1 cores takes ~ 2h for all genes
cds_pr_test_res <- graph_test(cds2, neighbor_graph="principal_graph", cores = 1)
cds_pr_test_res$symbol <- row.names(cds_pr_test_res)

# Note:  https://github.com/cole-trapnell-lab/monocle3/issues/509
# Need monocle3 dev version (use rbind instead of rBind in master branch)
# After installing dev version, seurat-wrapper converting seurat to cds doesn't work
# Have to install temp patch from: https://github.com/satijalab/seurat-wrappers/issues/119

```

```{r}
pr_deg_ids <- row.names(subset(cds_pr_test_res, q_value < 0.05))
```


Save DE result
```{r}
saveRDS(cds_pr_test_res, "data/20220404_graph_test_res.rds")
write.table(cds_pr_test_res, file = "data/20220404_graph_test_res.csv", quote = FALSE, sep = ",", row.names = F, col.names = T)
```



---

## Use harmony_SCT-integrated latent space

```{r}
All.rna.flt.dup <- All.rna.flt

All.rna.flt.dup@reductions[["umap"]] <- All.rna.flt.dup@reductions[["umap_harmony_SCT"]]
All.rna.flt.dup@reductions[["pca"]] <- All.rna.flt.dup@reductions[["harmony_SCT"]]

cds3 <- as.cell_data_set(All.rna.flt.dup)
cds3 <- cluster_cells(cds3)
plot_cells(cds3, color_cells_by = "partition", show_trajectory_graph = FALSE)

cds3 <- learn_graph(cds3, use_partition = TRUE)
plot_cells(cds3, label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = FALSE)


```


```{r}
cds3 <- learn_graph(cds3, use_partition = FALSE)
plot_cells(cds3, label_groups_by_cluster = FALSE, label_leaves = FALSE, label_branch_points = FALSE)
```





```{r}
cds3 <- order_cells(cds3)
plot_cells(cds3, color_cells_by = "pseudotime", label_cell_groups = FALSE, label_leaves = FALSE, label_branch_points = FALSE)
```


```{r}
tmp <- as.Seurat(cds3, data = NULL, assay = "data")

principal_graph_aux(cds3)$UMAP$pseudotime
```







---

# Other dimension reduction method for emphasizing continuous trajectories

---

## Force directed graph


Load reticulate and python env
```{r}

library(reticulate)
reticulate::use_condaenv("r-reticulate")

anndata = import("anndata", convert=FALSE)
#bbknn = import("bbknn", convert=FALSE)
sc = import("scanpy", convert=FALSE)
sc$settings$verbosity = 'debug' # ['error', 1:'warning', 2:'info', 3:'hint', 4:'debug']
sce = import("scanpy.external", convert=FALSE)

```




```{r}

count.raw <- Matrix::t(All.rna.flt@assays$RNA@counts) # transpose count for scanpy
varname <- data.frame(name=colnames(count.raw))
row.names(varname) <- varname$name

adata = anndata$AnnData(X=count.raw, obs=All.rna.flt@meta.data, var = varname)

sc$pp$normalize_total(adata, target_sum=1e4, exclude_highly_expressed='True')
sc$pp$log1p(adata)
sc$pp$highly_variable_genes(adata, n_top_genes=as.integer(2000), batch_key='sampleID')


hvg = py_to_r(adata$var$highly_variable)
adata.mat = py_to_r(adata$X)


```


```{r}
bdata = anndata$AnnData(X=adata.mat[,as.vector(hvg)], obs=All.rna.flt@meta.data)
sc$pp$regress_out(bdata, c('nCount_RNA', 'percent.mt'))
sc$pp$scale(bdata, max_value=as.integer(10))
sc$pp$pca(bdata, n_comps=as.integer(50), svd_solver='arpack')

sc$pp$neighbors(bdata)
sc$tl$draw_graph(bdata, layout='fa')
```



```{r}
sc$pl$draw_graph(bdata, color='sampleID')
```

```{r}
bdata
```

### Use bbknn integrated graph

```{r}
sce$pp$bbknn(bdata, batch_key="sampleID") 
```


```{r}
bdata
```


```{r}
sc$tl$draw_graph(bdata)
```

```{r}
sc$pl$draw_graph(bdata, color='sampleID')
```



---




```{r}
sc$tl$umap(bdata) 

```



```{r}
sc$pl$umap(bdata, color='sampleID') 

```







---

## Phate

---



```{r fig.height=5, fig.width=5}
Sys.time()

count.raw <- Matrix::t(seu@assays$RNA@counts) # transpose count for scanpy
varname <- data.frame(name=colnames(count.raw))
row.names(varname) <- varname$name

adata = anndata$AnnData(X=count.raw, obs=seu@meta.data, var = varname)
sc$pp$normalize_total(adata, target_sum=1e4, exclude_highly_expressed='True')
sc$pp$log1p(adata)
sc$pp$highly_variable_genes(adata, n_top_genes=as.integer(2000), batch_key='sampleID')

# Subset hvg for scaling & pca
hvg = py_to_r(adata$var$highly_variable)
adata.mat = py_to_r(adata$X)
bdata = anndata$AnnData(X=adata.mat[,as.vector(hvg)], obs=seu@meta.data)
sc$pp$regress_out(bdata, c('nCount_RNA', 'percent.mt'))
sc$pp$scale(bdata, max_value=as.integer(10))
sc$pp$pca(bdata, n_comps=as.integer(50), svd_solver='arpack')
sce$pp$bbknn(bdata, batch_key="sampleID") 
sc$tl$umap(bdata) 
sc$tl$tsne(bdata)

sc$tl$leiden(bdata, resolution=0.8, key_added='leiden_res_0_8')
sc$tl$leiden(bdata, resolution=0.5, key_added='leiden_res_0_5')
sc$tl$leiden(bdata, resolution=0.3, key_added='leiden_res_0_3')
sc$tl$leiden(bdata, resolution=0.1, key_added='leiden_res_0_1')


# Save umap coordinate and clustering back to seurat obj
umap = py_to_r(bdata$obsm[["X_umap"]])
row.names(umap) <- seu@reductions[["umap"]]@cell.embeddings %>% row.names
colnames(umap) <- c("umap_1", "umap_2")
seu@reductions[["umap_bbknn"]] <- seu@reductions[["umap"]]
seu@reductions[["umap_bbknn"]]@cell.embeddings <- umap
seu@reductions[["umap_bbknn"]]@key <- "umap_"
seu$bbknn_leiden_res_0_8 <- py_to_r(bdata$obs) %>% .$leiden_res_0_8
seu$bbknn_leiden_res_0_5 <- py_to_r(bdata$obs) %>% .$leiden_res_0_5
seu$bbknn_leiden_res_0_3 <- py_to_r(bdata$obs) %>% .$leiden_res_0_3
seu$bbknn_leiden_res_0_1 <- py_to_r(bdata$obs) %>% .$leiden_res_0_1


DimPlot(seu, group.by = "sampleID", cols = sampleID_color, raster=FALSE, shuffle=TRUE, reduction = "umap_bbknn") + ggtitle("Scanpy log1p + BBKNN") + NoAxes() 

Sys.time()

# Save tsne coordinate and clustering back to seurat obj
tsne = py_to_r(bdata$obsm[["X_tsne"]])
row.names(tsne) <- seu@reductions[["umap"]]@cell.embeddings %>% row.names
colnames(tsne) <- c("tsne_1", "tsne_2")
seu@reductions[["tsne_bbknn"]] <- seu@reductions[["umap"]]
seu@reductions[["tsne_bbknn"]]@cell.embeddings <- tsne
seu@reductions[["tsne_bbknn"]]@key <- "tsne_"

DimPlot(seu, group.by = "sampleID", cols = sampleID_color, raster=FALSE, shuffle=TRUE, reduction = "tsne_bbknn") + ggtitle("Scanpy standard + BBKNN") + NoAxes() 

```


---

# Session Info
```{r}
sessionInfo()
```

