---
title: "F6 interactome project: scRNA-seq analysis"
author: "Jen-Chien Chang"
date: "`r format(Sys.time(), '%Y/%m/%d')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, cache=FALSE, fig.show='hold', fig.align="center")
```

---

# Note

* 

---

# Basic setup

<span style="color:red">template code</span> <a href="#top">Back to top</a>
```{r}

```


Load R library
```{r warning=FALSE, results='hide'}
# Data manipulation
library(dplyr)
library(magrittr)
library(reshape2)

# Plot
library(ggplot2)
library(patchwork)
library(ggpubr)
library(rstatix)
library(pheatmap)
#library(ggfortify)    # for plot PCA with ggplot2 autoplot 

# File I/O
#library(readxl)

# Seurat
library(Seurat)
#library(SeuratDisk)

# Signac
#set.seed(2021)
#library(Signac)
#library(GenomeInfoDb)
#library(EnsDb.Mmusculus.v79)
#library(GenomicRanges)

# Multiprocessing
#library(future)
#plan("multiprocess", workers = 12)
#plan("sequential")

# Math
#library(usedist)

# For AUCell
#library(AUCell)
#library(GSEABase)

# sceasy (seurat <-> annData )
#library(sceasy)
#library(anndata) #directly deal with annData in R

```


Global variables
```{r}
# data directory
osc_fs_home_base <- "../../../../../osc-fs_home"
home_base <- "../../../../../home"

# color setting 
sampleID_color <- c("#e31a1c","#fb9a99","#33a02c","#b2df8a","#1f78b4","#a6cee3")
sampleType_color <- c("#e31a1c","#33a02c","#1f78b4")

# ggplot2 theme setting
theme_set(theme_bw()) 
```



---

# scRNA normalization, dimensional reduction, integration & clustering

---

Read filtered object
```{r}
seu <- readRDS("data/20220317_19535x26581_scRNA.rds")
```

---

## Standard Seurat workflow (log1p normalization to 1e4) without integration

```{r}

DefaultAssay(seu) <- "RNA"
  
Sys.time()

seu <- NormalizeData(seu)
seu <- FindVariableFeatures(seu)
seu <- ScaleData(seu)
seu <- RunPCA(seu)
seu <- RunUMAP(seu, dims = 1:30)
seu <- FindNeighbors(seu, dims = 1:30)
seu <- FindClusters(seu,  algorithm = 1 )
seu$log1p_Louvain_res.0.8 <-  seu$seurat_clusters
seu$RNA_snn_res.0.8 <- NULL
seu <- FindClusters(seu, algorithm = 4)
seu$log1p_Leiden_res.0.8 <-  seu$seurat_clusters
seu$RNA_snn_res.0.8 <- NULL
Sys.time() 



```

Perform CellCycleScoring
```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

seu <- CellCycleScoring(seu, g2m.features = g2m.genes, s.features = s.genes)
seu$Phase <- factor(seu$Phase)
```


```{r fig.height=4, fig.width=12}

p1 <- DimPlot(seu, group.by = "sampleID", cols = sampleID_color, raster=FALSE, shuffle=TRUE, reduction = "umap") + ggtitle("sampleID") + NoAxes()  
p2 <- DimPlot(seu, group.by = "log1p_Louvain_res.0.8", label = FALSE, raster=FALSE, reduction = "umap") + ggtitle("Louvain clustering") + NoAxes() + NoLegend()
p3 <- DimPlot(seu, group.by = "log1p_Leiden_res.0.8", label = FALSE, raster=FALSE, reduction = "umap") + ggtitle("Leiden clustering") + NoAxes() + NoLegend()
p1|p2|p3

#(FeaturePlot(seu, slot = "counts",features = c("Ttr"), pt.size=0.02) & NoAxes()) & scale_color_viridis_c(direction = 1)

```



---

## SCTransform workflow without integration

---

```{r}
Sys.time()
seu <- SCTransform(seu, method = "glmGamPoi")

Sys.time()
seu <- RunPCA(seu, reduction.name = "pca_SCT")
seu <- RunUMAP(seu, dims = 1:30, reduction = "pca_SCT", reduction.name = "umap_SCT")
seu <- FindNeighbors(seu, dims = 1:30, reduction = "pca_SCT")
seu <- FindClusters(seu)
Sys.time()

```



```{r fig.height=6, fig.width=12}

p1 <- DimPlot(seu, group.by = "sampleID", cols = sampleID_color, raster=FALSE, shuffle=TRUE, reduction = "umap_SCT") + ggtitle("tissue") + NoAxes()  
p2 <- DimPlot(seu, group.by = "SCT_snn_res.0.8", label = FALSE, raster=FALSE, reduction = "umap_SCT") + ggtitle("clusters") + NoAxes() + NoLegend()

p1|p2

```



---

## log1p normalization + RPCA integration

```{r}

Sys.time()
seu.list <- SplitObject(seu, split.by = "sampleID")

# normalize and identify variable features for each dataset independently
seu.list <- lapply(X = seu.list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

# select features that are repeatedly variable across datasets for integration run PCA on each
# dataset using these features
features <- SelectIntegrationFeatures(object.list = seu.list)
seu.list <- lapply(X = seu.list, FUN = function(x) {
    x <- ScaleData(x, features = features)
    x <- RunPCA(x, features = features)
})

# Find anchor
Sys.time()
seu.anchors <- FindIntegrationAnchors(object.list = seu.list, anchor.features = features, dims = 1:30, reduction = "rpca")

Sys.time()
seu.combined <- IntegrateData(anchorset = seu.anchors, dims = 1:30)

Sys.time()
```


```{r}
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
DefaultAssay(seu.combined) <- "integrated"

# Run the standard workflow for visualization and clustering
seu.combined <- ScaleData(seu.combined)

seu.combined <- RunPCA(seu.combined, npcs = 30)
seu.combined <- RunUMAP(seu.combined, reduction = "pca", dims = 1:30)

seu.combined <- FindNeighbors(seu.combined, dims = 1:30)
seu.combined <- FindClusters(seu.combined)
```


Plot on integrated umap
```{r fig.height=6, fig.width=12}

p1 <- DimPlot(seu.combined, group.by = "sampleID", cols = sampleID_color, raster=FALSE, shuffle=TRUE, reduction = "umap") + ggtitle("sampleID") + NoAxes()  
p2 <- DimPlot(seu.combined, group.by = "integrated_snn_res.0.8", label = FALSE, raster=FALSE, reduction = "umap") + ggtitle("clusters") + NoAxes() + NoLegend()

p1|p2

```


Save umap in the original object 
```{r}
seu@reductions[["umap_rpca"]] <- seu@reductions[["umap"]]
seu@reductions[["umap_rpca"]]@cell.embeddings <- seu.combined@reductions[["umap"]]@cell.embeddings
seu@reductions[["umap_rpca"]]@key <- "umap_"

seu$log1p_rpca_Louvain_res.0.8 <-  seu.combined$integrated_snn_res.0.8
```


save rpca combined object
```{r eval=FALSE}
saveRDS(seu.combined, "data/20220317_2000x26581_seu.combined.Rds")

rm(seu.list)
rm(seu.anchors)
rm(seu.combined)
```




---

## Integration using SCT & RPCA

```{r}

Sys.time()
seu.list <- SplitObject(seu, split.by = "sampleID")
seu.list <- lapply(X = seu.list, FUN = SCTransform, method = "glmGamPoi")

Sys.time()
features <- SelectIntegrationFeatures(object.list = seu.list, nfeatures = 3000)
seu.list <- PrepSCTIntegration(object.list = seu.list, anchor.features = features)
seu.list <- lapply(X = seu.list, FUN = RunPCA, features = features)

Sys.time()
seu.anchors <- FindIntegrationAnchors(object.list = seu.list, normalization.method = "SCT", anchor.features = features, dims = 1:30, reduction = "rpca")

Sys.time()
seu.combined.sct <- IntegrateData(anchorset = seu.anchors, normalization.method = "SCT", dims = 1:30)

Sys.time()
seu.combined.sct <- RunPCA(seu.combined.sct, verbose = FALSE)
seu.combined.sct <- RunUMAP(seu.combined.sct, reduction = "pca", dims = 1:30)

Sys.time()
seu.combined.sct <- FindNeighbors(seu.combined.sct, dims = 1:30)
seu.combined.sct <- FindClusters(seu.combined.sct)

Sys.time()
```


Plot
```{r fig.height=6, fig.width=12}

p1 <- DimPlot(seu.combined.sct, group.by = "sampleID", cols = sampleID_color, raster=FALSE, shuffle=TRUE, reduction = "umap") + ggtitle("sampleID") + NoAxes()  
p2 <- DimPlot(seu.combined.sct, group.by = "integrated_snn_res.0.8", label = FALSE, raster=FALSE, reduction = "umap") + ggtitle("clusters") + NoAxes() + NoLegend()

p1|p2

```


Save umap in the original object 
```{r}
seu@reductions[["umap_rpca_SCT"]] <- seu@reductions[["umap"]]
seu@reductions[["umap_rpca_SCT"]]@cell.embeddings <- seu.combined.sct@reductions[["umap"]]@cell.embeddings
seu@reductions[["umap_rpca_SCT"]]@key <- "umap_"

seu$SCT_rpca_Louvain_res.0.8 <-  seu.combined.sct$integrated_snn_res.0.8
```


save rpca combined object
```{r eval=FALSE}
saveRDS(seu.combined.sct, "data/20220317_3000x26581_seu.combined.sct.Rds")
rm(seu.list)
rm(seu.anchors)
rm(seu.combined.sct)
```



---

## log1p normalization + Harmony

Run Harmony 
```{r}
library(harmony)
DefaultAssay(seu) <- "RNA"

Sys.time()
seu <- RunHarmony(object = seu, group.by.vars = "sampleID" , plot_convergence = TRUE)
Sys.time()
DimPlot(object = seu, reduction = "harmony", pt.size = .1, group.by = "sampleID")
Sys.time()
seu <- RunUMAP(seu, reduction = "harmony", dims = 1:30, reduction.name = "umap_harmony" )
Sys.time()

seu <- FindNeighbors(seu, reduction = "harmony", dims = 1:30, graph.name = c("harmony_nn", "harmony_snn"))
seu <- FindClusters(seu, graph.name = "harmony_snn")

seu$harmony_Louvain_res.0.8 <- seu$harmony_snn_res.0.8
seu$harmony_snn_res.0.8 <- NULL
```


Plot
```{r fig.height=6, fig.width=12}

p1 <- DimPlot(seu, group.by = "sampleID", cols = sampleID_color, raster=FALSE, shuffle=TRUE, reduction = "umap_harmony") + ggtitle("sampleID") + NoAxes()  
p2 <- DimPlot(seu, group.by = "seurat_clusters", label = FALSE, raster=FALSE, reduction = "umap_harmony") + ggtitle("clusters") + NoAxes() + NoLegend()

p1|p2

```


---

## SCT + Harmony

```{r}
DefaultAssay(seu) <- "SCT"

Sys.time()
seu <- RunHarmony(object = seu, reduction = "pca_SCT", group.by.vars = "sampleID", plot_convergence = TRUE, reduction.save = "harmony_SCT", assay.use = "SCT")
Sys.time()
DimPlot(object = seu, reduction = "harmony_SCT", pt.size = .1, group.by = "sampleID")
Sys.time()
seu <- RunUMAP(seu, reduction = "harmony_SCT", dims = 1:30, reduction.name = "umap_harmony_SCT" )
Sys.time()
seu <- FindNeighbors(seu, reduction = "harmony_SCT", dims = 1:30, graph.name = c("harmony_SCT_nn", "harmony_SCT_snn"))
Sys.time()
seu <- FindClusters(seu, graph.name = "harmony_SCT_snn")
Sys.time()

seu$SCT_harmony_Louvain_res.0.8 <- seu$harmony_SCT_snn_res.0.8
seu$harmony_SCT_snn_res.0.8 <- NULL
```


```{r fig.height=6, fig.width=12}

p1 <- DimPlot(seu, group.by = "sampleID", cols = sampleID_color, raster=FALSE, shuffle=TRUE, reduction = "umap_harmony_SCT") + ggtitle("sampleID") + NoAxes()  
p2 <- DimPlot(seu, group.by = "seurat_clusters", label = FALSE, raster=FALSE, reduction = "umap_harmony_SCT") + ggtitle("clusters") + NoAxes() + NoLegend()

p1|p2

```


---

# Scanpy log1p normalization + BBKNN 


Load reticulate and python env
```{r}

library(reticulate)
reticulate::use_condaenv("r-reticulate")

anndata = import("anndata", convert=FALSE)
#bbknn = import("bbknn", convert=FALSE)
sc = import("scanpy", convert=FALSE)
sc$settings$verbosity = 'debug' # ['error', 1:'warning', 2:'info', 3:'hint', 4:'debug']
sce = import("scanpy.external", convert=FALSE)

```

## BBKNN integration
normalize to 1e4 (exclude highly expressed), HVG 2000 (batch = sampleID, flavor=1)
```{r fig.height=5, fig.width=5}
Sys.time()

count.raw <- Matrix::t(seu@assays$RNA@counts) # transpose count for scanpy
varname <- data.frame(name=colnames(count.raw))
row.names(varname) <- varname$name

adata = anndata$AnnData(X=count.raw, obs=seu@meta.data, var = varname)
sc$pp$normalize_total(adata, target_sum=1e4, exclude_highly_expressed='True')
sc$pp$log1p(adata)
sc$pp$highly_variable_genes(adata, n_top_genes=as.integer(2000), batch_key='sampleID')

# Subset hvg for scaling & pca
hvg = py_to_r(adata$var$highly_variable)
adata.mat = py_to_r(adata$X)
bdata = anndata$AnnData(X=adata.mat[,as.vector(hvg)], obs=seu@meta.data)
sc$pp$regress_out(bdata, c('nCount_RNA', 'percent.mt'))
sc$pp$scale(bdata, max_value=as.integer(10))
sc$pp$pca(bdata, n_comps=as.integer(50), svd_solver='arpack')
sce$pp$bbknn(bdata, batch_key="sampleID") 
sc$tl$umap(bdata) 
sc$tl$tsne(bdata)

sc$tl$leiden(bdata, resolution=0.8, key_added='leiden_res_0_8')
sc$tl$leiden(bdata, resolution=0.5, key_added='leiden_res_0_5')
sc$tl$leiden(bdata, resolution=0.3, key_added='leiden_res_0_3')
sc$tl$leiden(bdata, resolution=0.1, key_added='leiden_res_0_1')


# Save umap coordinate and clustering back to seurat obj
umap = py_to_r(bdata$obsm[["X_umap"]])
row.names(umap) <- seu@reductions[["umap"]]@cell.embeddings %>% row.names
colnames(umap) <- c("umap_1", "umap_2")
seu@reductions[["umap_bbknn"]] <- seu@reductions[["umap"]]
seu@reductions[["umap_bbknn"]]@cell.embeddings <- umap
seu@reductions[["umap_bbknn"]]@key <- "umap_"
seu$bbknn_leiden_res_0_8 <- py_to_r(bdata$obs) %>% .$leiden_res_0_8
seu$bbknn_leiden_res_0_5 <- py_to_r(bdata$obs) %>% .$leiden_res_0_5
seu$bbknn_leiden_res_0_3 <- py_to_r(bdata$obs) %>% .$leiden_res_0_3
seu$bbknn_leiden_res_0_1 <- py_to_r(bdata$obs) %>% .$leiden_res_0_1


DimPlot(seu, group.by = "sampleID", cols = sampleID_color, raster=FALSE, shuffle=TRUE, reduction = "umap_bbknn") + ggtitle("Scanpy log1p + BBKNN") + NoAxes() 

Sys.time()

# Save tsne coordinate and clustering back to seurat obj
tsne = py_to_r(bdata$obsm[["X_tsne"]])
row.names(tsne) <- seu@reductions[["umap"]]@cell.embeddings %>% row.names
colnames(tsne) <- c("tsne_1", "tsne_2")
seu@reductions[["tsne_bbknn"]] <- seu@reductions[["umap"]]
seu@reductions[["tsne_bbknn"]]@cell.embeddings <- tsne
seu@reductions[["tsne_bbknn"]]@key <- "tsne_"

DimPlot(seu, group.by = "sampleID", cols = sampleID_color, raster=FALSE, shuffle=TRUE, reduction = "tsne_bbknn") + ggtitle("Scanpy standard + BBKNN") + NoAxes() 

```



---

Save as h5ad for cellxgene (20220317)
```{r}
sceasy::convertFormat(seu, from="seurat", to="anndata", outFile='data/20220317_19535x26581_workInProgess.h5ad')
```

Save rds
```{r}
saveRDS(seu, "data/20220317_19535x26581_scRNA_processed.rds")
```



---

## Label transfer from Azimuth reference


### Human Fetus

Read ref rds
```{r}
seu.ref <- readRDS("../RefData/Azimuth/Human_Fetus/ref.Rds")
seu.ref$dataType <- "ref"

# 3000x377456
```

Plot ref data
```{r}
DimPlot(object = seu.ref, reduction = "refUMAP", group.by = "annotation.l1", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()
DimPlot(object = seu.ref, reduction = "refUMAP", group.by = "annotation.l2", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()

```


Find anchors 
```{r}
common.genes <- row.names(seu)[(row.names(seu) %in% row.names(seu.ref))]

anchors <- FindTransferAnchors(
  reference = seu.ref,
  features = common.genes, 
  query = seu,
  query.assay = "SCT",
  normalization.method = "SCT",
  reference.reduction = "refDR",
  dims = 1:50 
)
```


MapQuery
```{r}
seu.MQ <- MapQuery(
  anchorset = anchors,
  query = seu, 
  reference = seu.ref,
  refdata = list(
    annotation.l1 = "annotation.l1",
    annotation.l2 = "annotation.l2"
  ),
  reference.reduction = "refDR"
)
```

Seems ProjectUMAP is not run, run it
```{r}
seu.MQ <- ProjectUMAP(
  query = seu.MQ, 
  query.reduction = "ref.refDR", 
  reference = seu.ref, 
  reference.reduction = "refDR",
  reduction.model = "refUMAP"
)
```

Plot on ref umap
```{r fig.width=10}
DimPlot(object = seu.ref, reduction = "refUMAP", group.by = "annotation.l1") + NoLegend()
DimPlot(object = seu.MQ, reduction = "ref.umap", group.by = "predicted.annotation.l1", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()
```

Save predicted annotation back
```{r}
seu$LT_Azimuth_Fetus_l1 <- factor(seu.MQ$predicted.annotation.l1)
seu$LT_Azimuth_Fetus_l2 <- factor(seu.MQ$predicted.annotation.l2)
```

Plot on original umap
```{r}
DimPlot(object = seu, reduction = "umap_harmony", group.by = "LT_Azimuth_Fetus_l1", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()
DimPlot(object = seu, reduction = "umap_harmony", group.by = "LT_Azimuth_Fetus_l2", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()

```

save full LT result   
```{r}
saveRDS(seu.MQ, "data/20220318_seu.MQ_Fetus.Rds")
rm(seu.MQ)
rm(seu.ref)
```



---

### Human Motor Cortex

Read ref TMS rds
```{r}
seu.ref <- readRDS("../RefData/Azimuth/Human_MotorCortex/ref.Rds")
seu.ref$dataType <- "ref"

# 3000x24526
```

Plot ref data
```{r fig.width=10}
DimPlot(object = seu.ref, reduction = "refUMAP", group.by = "class", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()
DimPlot(object = seu.ref, reduction = "refUMAP", group.by = "subclass", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()
DimPlot(object = seu.ref, reduction = "refUMAP", group.by = "cross_species_cluster", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()
DimPlot(object = seu.ref, reduction = "refUMAP", group.by = "cluster", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()
```


Find anchors 
```{r}
common.genes <- row.names(seu)[(row.names(seu) %in% row.names(seu.ref))]

anchors <- FindTransferAnchors(
  reference = seu.ref,
  features = common.genes, 
  query = seu,
  query.assay = "SCT",
  normalization.method = "SCT",
  reference.reduction = "refDR",
  dims = 1:50 
)
```

MapQuery
```{r}
seu.MQ <- MapQuery(
  anchorset = anchors,
  query = seu, 
  reference = seu.ref,
  refdata = list(
    class = "class",
    subclass = "subclass"
  ),
  reference.reduction = "refDR"
)
```


Seems ProjectUMAP is not run, run it
```{r}
seu.MQ <- ProjectUMAP(
  query = seu.MQ, 
  query.reduction = "ref.refDR", 
  reference = seu.ref, 
  reference.reduction = "refDR",
  reduction.model = "refUMAP"
)
```

Plot on ref umap
```{r fig.width=10}
DimPlot(object = seu.ref, reduction = "refUMAP", group.by = "class") + NoLegend()
DimPlot(object = seu.MQ, reduction = "ref.umap", group.by = "predicted.class", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()

DimPlot(object = seu.ref, reduction = "refUMAP", group.by = "subclass") + NoLegend()
DimPlot(object = seu.MQ, reduction = "ref.umap", group.by = "predicted.subclass", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()
```

Save predicted annotation back
```{r}
seu$LT_Azimuth_MotorCortex_class <- factor(seu.MQ$predicted.class)
seu$LT_Azimuth_MotorCortex_subclass <- factor(seu.MQ$predicted.subclass)
```

Plot on original umap
```{r}
DimPlot(object = seu, reduction = "umap_harmony", group.by = "LT_Azimuth_MotorCortex_class", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()
DimPlot(object = seu, reduction = "umap_harmony", group.by = "LT_Azimuth_MotorCortex_subclass", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()

```

save full LT result   
```{r}
saveRDS(seu.MQ, "data/20220318_seu.MQ_MotorCortex.Rds")
rm(seu.MQ)
rm(seu.ref)
```


---

# Annotation using sc-type

Follow the [Github example](https://github.com/IanevskiAleksandr/sc-type)

Source the code
```{r}
# load gene set preparation function
source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/gene_sets_prepare.R")
# load cell type annotation function
source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/sctype_score_.R")
```

Prepare db
```{r}
# DB file
db_ = "https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/ScTypeDB_full.xlsx";
tissue = "Brain" # e.g. Immune system, Liver, Pancreas, Kidney, Eye, Brain

# prepare gene sets
library(HGNChelper)
gs_list = gene_sets_prepare(db_, tissue)
```

Example data
```{r}
gs_list.ex = gene_sets_prepare("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/ScTypeDB_short.xlsx", "Immune system") # e.g. Immune system, Liver, Pancreas, Kidney, Eye, Brain

# assign cell types
scRNAseqData = readRDS(gzcon(url('https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/exampleData.RDS'))); #load example scRNA-seq matrix
es.max = sctype_score(scRNAseqData = scRNAseqData, scaled = TRUE, gs = gs_list.ex$gs_positive, gs2 = gs_list.ex$gs_negative)

# View results, cell-type by cell matrix. See the complete example below
View(es.max)
```

Apply to our data
```{r}
# Use brain geneset in scType database
gs_list = gene_sets_prepare(db_, "Brain")

# get cell-type by cell matrix
es.max = sctype_score(scRNAseqData = seu[["RNA"]]@scale.data, scaled = TRUE, 
                      gs = gs_list$gs_positive, gs2 = gs_list$gs_negative)

# merge by cluster: harmony_Louvain_res.0.8
cL_resutls = do.call("rbind", lapply(unique(seu@meta.data$harmony_Louvain_res.0.8), function(cl){
    es.max.cl = sort(rowSums(es.max[ ,rownames(seu@meta.data[seu@meta.data$harmony_Louvain_res.0.8==cl, ])]), decreasing = !0)
    head(data.frame(cluster = cl, type = names(es.max.cl), scores = es.max.cl, ncells = sum(seu@meta.data$harmony_Louvain_res.0.8==cl)), 10)
}))
sctype_scores = cL_resutls %>% group_by(cluster) %>% top_n(n = 1, wt = scores)  

# set low-confident (low ScType score) clusters to "unknown"
sctype_scores$type[as.numeric(as.character(sctype_scores$scores)) < sctype_scores$ncells/4] = "Unknown"
DT::datatable(sctype_scores)
```

Plot scType assigned annotation on harmony_umap
```{r}
seu@meta.data$scType_assigned_harmony_Louvain_res.0.8 = ""
for(j in unique(sctype_scores$cluster)){
  cl_type = sctype_scores[sctype_scores$cluster==j,]; 
  seu@meta.data$scType_assigned_harmony_Louvain_res.0.8[seu@meta.data$harmony_Louvain_res.0.8 == j] = as.character(cl_type$type[1])
}
seu$scType_assigned_harmony_Louvain_res.0.8 <- factor(seu$scType_assigned_harmony_Louvain_res.0.8)

DimPlot(seu, reduction = "umap_harmony", label = TRUE, repel = TRUE, group.by = 'scType_assigned_harmony_Louvain_res.0.8') 

```


Save scType raw 
```{r}

seu$scType_raw <- factor(colnames(t(es.max))[max.col(t(es.max), ties.method="first")])

DimPlot(seu, reduction = "umap_harmony", group.by = 'scType_raw', label = T, label.size = 3, repel = TRUE) + NoLegend()

```





---


Save as h5ad for cellxgene (20220318)
```{r}
sceasy::convertFormat(seu, from="seurat", to="anndata", outFile='data/20220317_19535x26581_workInProgess.h5ad')
```

Save rds

```{r}
saveRDS(seu, "data/20220317_19535x26581_scRNA_processed.rds")
```



---

# Session Info
```{r}
sessionInfo()
```

