---
title: "F6 interactome project: scRNA-seq QC and filtering"
author: "Jen-Chien Chang"
date: "`r format(Sys.time(), '%Y/%m/%d')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, cache=FALSE, fig.show='hold', fig.align="center")
```

---

# Note

* 

---

# Basic setup

<span style="color:red">template code</span> <a href="#top">Back to top</a>
```{r}

```


Load R library
```{r warning=FALSE, results='hide'}
# Data manipulation
library(dplyr)
library(magrittr)
library(reshape2)

# Plot
library(ggplot2)
library(patchwork)
library(ggpubr)
library(rstatix)
library(pheatmap)
#library(ggfortify)    # for plot PCA with ggplot2 autoplot 

# File I/O
#library(readxl)

# Seurat
library(Seurat)
#library(SeuratDisk)

# Signac
#set.seed(2021)
#library(Signac)
#library(GenomeInfoDb)
#library(EnsDb.Mmusculus.v79)
#library(GenomicRanges)

# Multiprocessing
#library(future)
#plan("multiprocess", workers = 12)
#plan("sequential")

# Math
#library(usedist)

# For AUCell
#library(AUCell)
#library(GSEABase)

# sceasy (seurat <-> annData )
#library(sceasy)
#library(anndata) #directly deal with annData in R

```


Global variables
```{r}
# data directory
osc_fs_home_base <- "../../../../../osc-fs_home"
home_base <- "../../../../../home"

# color setting 
sampleID_color <- c("#e31a1c","#fb9a99","#33a02c","#b2df8a","#1f78b4","#a6cee3")
sampleType_color <- c("#e31a1c","#33a02c","#1f78b4")

# ggplot2 theme setting
theme_set(theme_bw()) 
```



---

# scRNA-seq analysis

---


## Cell QC & filtering


Read data with scrublet output back in R
```{r}
seu.raw <- readRDS('data/20220317_raw_merged_scrublet.rds')
seu.raw
seu.raw@meta.data %>% colnames
seu.raw$sampleID <- factor(seu.raw$sampleID, c("iPS1","iPS2","NSC1","NSC2","Neuron1","Neuron2"))
seu.raw$sampleType <- seu.raw$sampleID %>% as.character() %>% strsplit(., "[0-9]") %>% unlist
seu.raw$sampleType <- factor(seu.raw$sampleType, c("iPS","NSC","Neuron"))
```


### doublet removal

Plot doublet score
```{r}
seu.raw@meta.data %>%
  ggplot(., aes(x=nCount_RNA, y=doublet_scores))+
  geom_point(size = 0.5, stroke = 0, shape = 16)

seu.raw@meta.data %>%
  ggplot(., aes(x=doublet_scores, y=predicted_doublets))+
  geom_point(size = 0.5, stroke = 0, shape = 16)

seu.raw@meta.data %>%
  ggplot(., aes(x=nCount_RNA, y=predicted_doublets))+
  geom_point(size = 0.5, stroke = 0, shape = 16)
```

Remove cells labelled as doublet
```{r}

nDoublet <- nrow(filter(seu.raw@meta.data, predicted_doublets == TRUE))
print(paste0("Num. cells before all filtering: ", nrow(seu.raw@meta.data)))
print(paste0("Num. doublets predicted by scrublet: ", nDoublet, " (", round(nDoublet/nrow(seu.raw@meta.data)*100, 2),"%)"))
print(paste0("Num. cells after doublet removal: ", nrow(seu.raw@meta.data)-nDoublet))
```

### Calculate percentage of reads for various gene category 
```{r}

seu.raw[["percent.mt"]] <- PercentageFeatureSet(seu.raw, pattern = "^MT-")
seu.raw[["percent.ribo"]] <- PercentageFeatureSet(seu.raw, pattern = "^(RPL|RPL)")
seu.raw[["percent.hb"]] <- PercentageFeatureSet(seu.raw, pattern = "^HB")
seu.raw[["percent.hist"]] <- PercentageFeatureSet(seu.raw, pattern = "^(HIST|H1F|H2A|H2B|H3|H4)")


print(paste0("Mitochondrial genes: ", length(grep("^MT-", rownames(seu.raw@assays$RNA@counts), value = TRUE)) ))
grep("^MT-", rownames(seu.raw@assays$RNA@counts), value = TRUE)

print(paste0("Ribosomal genes: ", length(grep("^(RPL|RPL)", rownames(seu.raw@assays$RNA@counts), value = TRUE))))
grep("^(RPL|RPL)", rownames(seu.raw@assays$RNA@counts), value = TRUE)

print(paste0("Hemoglobin genes: ", length(grep("^HB", rownames(seu.raw@assays$RNA@counts), value = TRUE))))
grep("^HB", rownames(seu.raw@assays$RNA@counts), value = TRUE)

print(paste0("Histone genes: ", length(grep("^(HIST|H1F|H2A|H2B|H3|H4)", rownames(seu.raw@assays$RNA@counts), value = TRUE))))
grep("^(HIST|H1F|H2A|H2B|H3|H4)", rownames(seu.raw@assays$RNA@counts), value = TRUE)

```



### Plot quality metrics

```{r fig.}
VlnPlot(seu.raw, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = "sampleID", pt.size = 0)
```


```{r}
seu.raw@meta.data %>%
  ggplot(., aes(x=log2(nCount_RNA), y=log2(nFeature_RNA), color=percent.mt))+
  geom_point(size = 0.1, stroke = 0, shape = 16)+
  scale_color_viridis_c(direction = 1, limits = c(0, 100))+
  ylim(c(6,14))+ xlim(c(8,17))+
  geom_vline(aes(xintercept=log2(500)), color="black", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=log2(1000)), color="red", linetype="dashed", size=1)+  
  geom_hline(aes(yintercept=log2(250)), color="black", linetype="dashed", size=1)+
  geom_hline(aes(yintercept=log2(500)), color="red", linetype="dashed", size=1)+
  ggtitle("All cells")
```



```{r}
df <- seu.raw@meta.data

p0 <- df %>% group_by(sampleID) %>% mutate(., nCells=n()) %>% distinct(sampleID, .keep_all = T) %>% 
  ggplot(., aes(x=nCells, y=sampleID, color=sampleID))+
  geom_point(position = position_dodge(width = 0.5))+
  scale_color_manual(values=sampleID_color)+
  theme(legend.position = "none", axis.title.y=element_blank(), axis.text.y=element_blank())


## nCount_RNA (nUMI)

print(paste0("nCount_RNA (nUMI): "))
quantile(df$nCount_RNA, c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=nCount_RNA)) +
    geom_histogram(binwidth=100, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(nCount_RNA, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(nCount_RNA, na.rm=T)), color="black", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=100), color="blue", size=1)+
  geom_vline(aes(xintercept=25000), color="blue", size=1)+
  xlim(c(0,50000))+
  ggtitle("nCount_RNA (nUMI)")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=nCount_RNA, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,50000))+
  theme(axis.title.y=element_blank())
p3 <- ggplot(df, aes(x=nCount_RNA, y=all))+
  geom_boxplot()+
  xlim(c(0,50000))+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))
#p1+p3+p2+plot_layout(heights = c(1,0.3,2))


## nFeature_RNA (nGenes)

print(paste0("nFeature_RNA (nGenes): "))
df$nFeature_RNA %>% quantile(., c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=nFeature_RNA)) +
    geom_histogram(binwidth=50, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(nFeature_RNA, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(nFeature_RNA, na.rm=T)), color="black", linetype="dashed", size=1)+
  xlim(c(0,10000))+
  ggtitle("nFeature_RNA (nGenes)")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=nFeature_RNA, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,10000))+
  theme(axis.title.y=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))

## Percentage of Mitochondrial Reads

print(paste0("Percentage of Mitochondrial Reads: "))
df$percent.mt %>% quantile(., c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=percent.mt)) +
    geom_histogram(binwidth=0.5, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(percent.mt, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(percent.mt, na.rm=T)), color="black", linetype="dashed", size=1)+
  xlim(c(0,30))+
  ggtitle("Percentage of Mitochondrial Reads")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=percent.mt, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,30))+
  theme(axis.title.y=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))

## Percentage of Ribosomal Reads

print(paste0("Percentage of Ribosomal Reads: "))
df$percent.ribo %>% quantile(., c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=percent.ribo)) +
    geom_histogram(binwidth=0.5, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(percent.ribo, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(percent.ribo, na.rm=T)), color="black", linetype="dashed", size=1)+
  xlim(c(0,30))+
  ggtitle("Percentage of Ribosomal Reads")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=percent.ribo, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,30))+
  theme(axis.title.y=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))



## Percentage of Hb Reads

print(paste0("Percentage of Hemoglobin Reads: "))
df$percent.hb %>% quantile(., c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=percent.hb)) +
    geom_histogram(binwidth=0.05, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(percent.hb, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(percent.hb, na.rm=T)), color="black", linetype="dashed", size=1)+
  xlim(c(0,5))+
  ggtitle("Percentage of Hemoglobin Reads")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=percent.hb, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,5))+
  theme(axis.title.y=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))



## Percentage of Histone Reads

print(paste0("Percentage of Histone Reads: "))
df$percent.hist %>% quantile(., c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=percent.hist)) +
    geom_histogram(binwidth=0.05, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(percent.hist, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(percent.hist, na.rm=T)), color="black", linetype="dashed", size=1)+
  xlim(c(0,3))+
  ggtitle("Percentage of Histone Reads")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=percent.hist, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,3))+
  theme(axis.title.y=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))


## doublet score distribution
df %>%
  ggplot(., aes(x=doublet_scores)) +
    geom_histogram(binwidth=0.01, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(doublet_scores, na.rm=T)), color="red", linetype="dashed", size=1)+
    geom_vline(aes(xintercept=(mean(doublet_scores, na.rm=T)+3*sd(doublet_scores, na.rm=T))), color="red", linetype="dashed", size=1)+
    geom_vline(aes(xintercept=median(doublet_scores, na.rm=T)), color="blue", linetype="dashed", size=1)+
  ggtitle("Doublet score distribution")


## pct.mt vs pct.ribo
ggplot(df, aes(x=percent.ribo, y=percent.mt))+
  geom_point()+
  ggtitle("pct.mt vs pct.ribo")

```

---

### Apply filtering 

Set filtering criteria
```{r}
minUMI <- 2000
minGenes <- 1000
maxPctMt <- 20

seu <- subset(seu.raw, subset = (nCount_RNA > minUMI & nFeature_RNA > minGenes & percent.mt < maxPctMt & predicted_doublets == FALSE))

df <- seu@meta.data

nFlt <- nrow(seu.raw@meta.data) - nrow(df)
print(paste0("Num. cells before all filtering: ", nrow(seu.raw@meta.data)))
print(paste0("Num. filtered: ", nFlt, " (", round(nFlt/nrow(seu.raw@meta.data)*100, 2),"%)"))
print(paste0("Num. cells after all filtering: ", nrow(df)))

```

Stats after all filtering
```{r}
  
p0 <- df %>% group_by(sampleID) %>% mutate(., nCells=n()) %>% distinct(sampleID, .keep_all = T) %>% 
  ggplot(., aes(x=nCells, y=sampleID, color=sampleID))+
  geom_point(position = position_dodge(width = 0.5))+
  scale_color_manual(values=sampleID_color)+
  theme(legend.position = "none", axis.title.y=element_blank(), axis.text.y=element_blank())

## log2(nUMI) vs log2(nGenes)

ggplot(df, aes(x=log2(nCount_RNA), y=log2(nFeature_RNA), color=percent.mt))+
            geom_point(data = filter(seu.raw@meta.data, !(row.names(seu.raw@meta.data) %in% row.names(df))), size = 0.1, stroke = 0, shape = 16, color="grey")+ 
            geom_point(size = 0.1, stroke = 0, shape = 16)+ 
            scale_color_viridis_c(direction = 1, limits = c(0, 50))+
            ylim(c(6,13))+ xlim(c(8,17))+
            geom_vline(aes(xintercept=log2(500)), color="black", linetype="dashed", size=1)+
            geom_vline(aes(xintercept=log2(1000)), color="red", linetype="dashed", size=1)+  
            geom_hline(aes(yintercept=log2(250)), color="black", linetype="dashed", size=1)+
            geom_hline(aes(yintercept=log2(500)), color="red", linetype="dashed", size=1)+
            #geom_abline(slope = fit()[2] , intercept = fit()[4]  , color="blue", linetype="dashed", size=1)+
            #geom_abline(slope = fit()[2] , intercept = fit()[3]  , color="blue", linetype="dashed", size=1)+
            #geom_abline(slope = fit()[2] , intercept = fit()[1] , color="blue", linetype="dashed", size=1)+
            ggtitle(paste0("nCells=", nrow(df), " (", nrow(seu.raw@meta.data)-nrow(df), " removed, ", round((nrow(seu.raw@meta.data)-nrow(df))/nrow(seu.raw@meta.data)*100, 1), "%)"))+ 
            theme(plot.title = element_text(size = 12, face = "bold"))


## nCount_RNA (nUMI)

print(paste0("nCount_RNA (nUMI): "))
quantile(df$nCount_RNA, c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=nCount_RNA)) +
    geom_histogram(binwidth=100, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(nCount_RNA, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(nCount_RNA, na.rm=T)), color="black", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=100), color="blue", size=1)+
  geom_vline(aes(xintercept=25000), color="blue", size=1)+
  xlim(c(0,50000))+
  ggtitle("nCount_RNA (nUMI)")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=nCount_RNA, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,50000))+
  theme(axis.title.y=element_blank())
p3 <- ggplot(df, aes(x=nCount_RNA, y=all))+
  geom_boxplot()+
  xlim(c(0,50000))+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))
#p1+p3+p2+plot_layout(heights = c(1,0.3,2))


## nFeature_RNA (nGenes)

print(paste0("nFeature_RNA (nGenes): "))
df$nFeature_RNA %>% quantile(., c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=nFeature_RNA)) +
    geom_histogram(binwidth=50, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(nFeature_RNA, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(nFeature_RNA, na.rm=T)), color="black", linetype="dashed", size=1)+
  xlim(c(0,10000))+
  ggtitle("nFeature_RNA (nGenes)")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=nFeature_RNA, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,10000))+
  theme(axis.title.y=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))

## Percentage of Mitochondrial Reads

print(paste0("Percentage of Mitochondrial Reads: "))
df$percent.mt %>% quantile(., c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=percent.mt)) +
    geom_histogram(binwidth=0.5, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(percent.mt, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(percent.mt, na.rm=T)), color="black", linetype="dashed", size=1)+
  xlim(c(0,25))+
  ggtitle("Percentage of Mitochondrial Reads")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=percent.mt, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,25))+
  theme(axis.title.y=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))

## Percentage of Ribosomal Reads

print(paste0("Percentage of Ribosomal Reads: "))
df$percent.ribo %>% quantile(., c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=percent.ribo)) +
    geom_histogram(binwidth=0.5, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(percent.ribo, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(percent.ribo, na.rm=T)), color="black", linetype="dashed", size=1)+
  xlim(c(0,30))+
  ggtitle("Percentage of Ribosomal Reads")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=percent.ribo, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,30))+
  theme(axis.title.y=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))



## Percentage of Hb Reads

print(paste0("Percentage of Hemoglobin Reads: "))
df$percent.hb %>% quantile(., c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=percent.hb)) +
    geom_histogram(binwidth=0.05, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(percent.hb, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(percent.hb, na.rm=T)), color="black", linetype="dashed", size=1)+
  xlim(c(0,5))+
  ggtitle("Percentage of Hemoglobin Reads")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=percent.hb, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,5))+
  theme(axis.title.y=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))



## Percentage of Histone Reads

print(paste0("Percentage of Histone Reads: "))
df$percent.hist %>% quantile(., c(0,0.01,0.05,0.25,0.5,0.75,0.95,0.99,1))

p1 <- ggplot(df, aes(x=percent.hist)) +
    geom_histogram(binwidth=0.05, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(percent.hist, na.rm=T)), color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=median(percent.hist, na.rm=T)), color="black", linetype="dashed", size=1)+
  xlim(c(0,3))+
  ggtitle("Percentage of Histone Reads")+
  theme(axis.title=element_blank(), axis.text.x=element_blank())
p2 <- ggplot(df, aes(x=percent.hist, y=sampleID))+
  geom_boxplot()+
  xlim(c(0,3))+
  theme(axis.title.y=element_blank())
p1+plot_spacer()+p2+p0 + plot_layout(widths = c(2, 1), heights = c(1,2))


## doublet score distribution
df %>%
  ggplot(., aes(x=doublet_scores)) +
    geom_histogram(binwidth=0.01, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(doublet_scores, na.rm=T)), color="red", linetype="dashed", size=1)+
    geom_vline(aes(xintercept=(mean(doublet_scores, na.rm=T)+3*sd(doublet_scores, na.rm=T))), color="red", linetype="dashed", size=1)+
    geom_vline(aes(xintercept=median(doublet_scores, na.rm=T)), color="blue", linetype="dashed", size=1)+
  ggtitle("Doublet score distribution")


## pct.mt vs pct.ribo
ggplot(df, aes(x=percent.ribo, y=percent.mt))+
  geom_point()+
  ggtitle("pct.mt vs pct.ribo")

```




## Further filter genes  

Keep only genes expressing in >50 cells
```{r}
min.cell <- 50
nCellsExpr <- Matrix::rowSums(seu@assays$RNA@counts > 0)
 
seu <- seu[names(nCellsExpr[nCellsExpr > min.cell]),]
seu

rm(seu.raw)
```

save rds
```{r eval=FALSE}

saveRDS(seu, "data/20220317_19535x26581_scRNA.rds")

```



---

# Session Info
```{r}
sessionInfo()
```

