---
title: "F6 interactome project: scRNA-seq processing and filtering"
author: "Jen-Chien Chang"
date: "`r format(Sys.time(), '%Y/%m/%d')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, cache=FALSE, fig.show='hold', fig.align="center")
```

---

# Note

* 

---

# Basic setup

<span style="color:red">template code</span> <a href="#top">Back to top</a>
```{r}

```


Load R library
```{r warning=FALSE, results='hide'}
# Data manipulation
library(dplyr)
library(magrittr)
library(reshape2)

# Plot
library(ggplot2)
library(patchwork)
library(ggpubr)
library(rstatix)
library(pheatmap)
#library(ggfortify)    # for plot PCA with ggplot2 autoplot 

# File I/O
#library(readxl)

# Seurat
library(Seurat)
#library(SeuratDisk)

# Signac
#set.seed(2021)
#library(Signac)
#library(GenomeInfoDb)
#library(EnsDb.Mmusculus.v79)
#library(GenomicRanges)

# Multiprocessing
#library(future)
#plan("multiprocess", workers = 12)
#plan("sequential")

# Math
#library(usedist)

# For AUCell
#library(AUCell)
#library(GSEABase)

# sceasy (seurat <-> annData )
#library(sceasy)
#library(anndata) #directly deal with annData in R

```


Global variables
```{r}
# data directory
osc_fs_home_base <- "../../../../../osc-fs_home"
home_base <- "../../../../../home"

# color setting 
sampleID_color <- c("#e31a1c","#fb9a99","#33a02c","#b2df8a","#1f78b4","#a6cee3")
sampleType_color <- c("#e31a1c","#33a02c","#1f78b4")

# ggplot2 theme setting
theme_set(theme_bw()) 
```



---

# scRNA-seq analysis

---

## Read cellranger output

Read Cellranger output for 6 samples 
```{r eval=FALSE}
seu.list <- list()

for(i in c("iPS1","iPS2","Neuron1","Neuron2","NSC1","NSC2")){

tmp <- Read10X(data.dir = paste0("../F6_Interactome_20220128/scRNA/",i,"/outs/filtered_feature_bc_matrix/"))
tmp <- CreateSeuratObject(counts = tmp, project = i, min.cells = 0, min.features = 0)
seu.list[[i]] <- tmp

}

```

Concatenate all samples. 
```{r eval=FALSE}
tmp<-c()
for(i in 2:length(seu.list)){
  tmp <- c(tmp, seu.list[[i]])
}
seu.all <- merge(seu.list[[1]], y = tmp) # "_1" ~ "_6" added to the end of original cellbarcodes
rm(tmp)
```


```{r eval=FALSE}
seu.all$sampleID <- factor(seu.all$orig.ident, c("iPS1","iPS2","NSC1","NSC2","Neuron1","Neuron2"))
```


convert to h5ad for scanpy
```{r eval=FALSE}
sceasy::convertFormat(seu.all, from="seurat", to="anndata", outFile='data/20220317_raw_merged.h5ad')
```


---

## Run Scrublet to identify potential doublet


```{r}
# Run python in Rmd
library(reticulate)
reticulate::use_condaenv("r-reticulate")
```

Import python package
```{python, error=TRUE, eval=FALSE}
import matplotlib.pyplot as pl
import seaborn as sns
import numpy as np
import pandas as pd
import scanpy as sc
sc.settings.verbosity = 3

from matplotlib.pyplot import rc_context
import scrublet as scr
import anndata as ad
from datetime import datetime

import skimage
```

Read raw h5ad and split 
```{python, error=TRUE, eval=FALSE}
adata = sc.read('data/20220317_raw_merged.h5ad')

# split by sampleID
groups = adata.obs.groupby("sampleID").indices
adataList = [adata[inds] for inds in groups.values()]
```

Run scrublet individually
```{python, error=TRUE, eval=FALSE)}
print(datetime.now().time())

# Run Scrublet on 145 individual adata
for i in range(len(adataList)):
  
  tmp_adata = adataList[i]
  print('\n ###### No.', i, ': sampleID =', adataList[i].obs['sampleID'][0])
  # Run scrublet with expected doublet rate set to 5% (6000 cells)
  scrub = scr.Scrublet(tmp_adata.X, expected_doublet_rate=0.05)
  tmp_adata.obs['doublet_scores'], tmp_adata.obs['predicted_doublets_min'] = scrub.scrub_doublets()
  # Find the new threshold by using threshold_otsu
  th_orig = scrub.threshold_
  print("  => Original threshold by Minimum Method at doublet score = {:.2f}".format(th_orig))
  th_otsu = skimage.filters.threshold_otsu(image=scrub.doublet_scores_sim_)
  print("  => New threshold by Otsu Method at doublet score = {:.2f}".format(th_otsu))
  # Call doublet again using new threshold
  tmp_adata.obs['predicted_doublets'] = scrub.call_doublets(threshold=th_otsu)
  # Plot the histogram to check
  hist = scrub.plot_histogram()
  hist[0].suptitle(adataList[i].obs['sampleID'][0], x=0.1, y=0.98)
  ax = hist[1][0]
  yl = ax.get_ylim()
  ax.plot(th_orig * np.ones(2), yl, c = 'blue', linewidth = 1)  
  ax = hist[1][1]
  yl = ax.get_ylim()
  ax.plot(th_orig * np.ones(2), yl, c = 'blue', linewidth = 1)  
  pl.show()
  # Save data back
  adataList[i] = tmp_adata

print(datetime.now().time())

```


Concat all adata
```{python, error=TRUE, eval=FALSE}
adata_concat = ad.concat(adataList, merge="same")
adata_concat.write('data/20220317_raw_merged_scrublet.h5ad')
print(datetime.now().time())
```


Convert h5ad to rds
```{r eval=FALSE}
sceasy::convertFormat('data/20220317_raw_merged_scrublet.h5ad', from="anndata", to="seurat",
                      outFile='data/20220317_raw_merged_scrublet.rds')
```



---

# Session Info
```{r}
sessionInfo()
```

